---
title: "ALS Processing Benchmark Report"
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: false
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load}
#| include: false
source("R/report-helpers.R")
library(gt)

results <- load_all_benchmarks()

# What each benchmark tests and why, keyed by its .RDS basename (see
# benchmarks/*.R for the code these describe). Shown above each table below.
# Human-readable title for each benchmark, shown as its section heading
# instead of the raw .RDS/script basename.
benchmark_titles <- c(
  tools_rw = "Read and Write",
  tools_copc = "Convert to COPC",
  tools_normals = "Compute Normals",
  tools_ground = "Classify Ground",
  lasr_stages = "lasR Pipeline Stages",
  `lidr-lasr_ground-intensity` = "Ground Classification + Intensity Normalization",
  parallel_strategies = "Parallelization Strategies",
  `drives_local-vs-network` = "Local vs. Network Drive",
  `workers_raw-to-processed` = "Worker-Count Scaling"
)

display_title <- function(b) {
  if (b %in% names(benchmark_titles)) benchmark_titles[[b]] else b
}

benchmark_context <- c(
  tools_rw = "Reads a single sample file and
    writes it straight back out again, once with each tool (lidR, lasR,
    PDAL, LAStools). lidR and lasR are limited to 4 CPU cores so neither
    gets an unfair speed advantage over the other; PDAL and LAStools run
    with their own default, unrestricted parallelism, since they're
    invoked as external command-line tools with no core limit applied.
    This is the baseline read/write cost that every other tool
    comparison builds on.",
  tools_copc = "Converts a single sample file into Cloud-Optimized Point
    Cloud (COPC) format, once with lasR, PDAL, and LAStools (lidR can't
    write COPC). lasR is limited to 4 CPU cores; PDAL and LAStools run
    with their own default, unrestricted parallelism, since they're
    invoked as external command-line tools with no core limit applied.
    COPC is a format we're considering as our delivery format, so this
    is about its conversion cost.",
  tools_normals = "Computes surface normals (a per-point geometry
    feature) for a single sample file, once with lasR and once with PDAL
    (lidR and LAStools don't offer this). lasR is limited to 4 CPU
    cores; PDAL runs with its own default, unrestricted parallelism,
    since it's invoked as an external command-line tool with no core
    limit applied. Relevant to any pipeline that needs geometric point
    features downstream.",
  tools_ground = "Classifies ground points in a single sample file, once
    per tool's ground-classification method (lidR/lasR/PDAL's CSF
    algorithm, lasR's own alternative method, LAStools' own algorithm).
    lidR and lasR are limited to 4 CPU cores; PDAL and LAStools run with
    their own default, unrestricted parallelism, since they're invoked
    as external command-line tools with no core limit applied. Ground
    classification has to happen before terrain models can be built, so
    its cost matters at full scale. Note: these methods are not
    equivalent — different algorithms and parameters will classify some
    points differently, so this is a speed comparison, not a proof that
    the outputs match.",
  lasr_stages = "Times individual lasR stages — reading,
    writing in three formats, sorting, building a terrain surface,
    rasterizing, generating a ground or surface elevation model,
    normalizing point heights, finding tree tops, and four different
    ground/noise classification methods — on a single sample file
    already loaded into memory, limited to 4 cores. Shows which steps
    are the expensive ones in a typical processing pipeline.",
  `lidr-lasr_ground-intensity` = "Runs a realistic two-step pipeline —
    classify ground points, then rescale intensity values — on a single
    sample file, once written as a pipeline in lidR and once as a lasR
    pipeline with that same custom R logic plugged in. Same 4-core limit.
    Tests whether lasR stays faster than lidR once real custom logic
    (not just lasR's built-in steps) is part of the pipeline.",
  parallel_strategies = "Reads and writes all 100 sample files from local
    storage, using three different ways of spreading that work across
    CPU cores. Each strategy uses half of the machine's available
    cores. Shows which of the three approaches production tooling should
    be built on. Caveat: the lasr_internal method hands lasR the entire
    set of 100 files in one call, which makes lasR read extra data
    around each file's edges to stay aware of its neighboring tiles —
    work the other two methods skip entirely, since each hands lasR one
    completely isolated file at a time. So lasr_internal being slower
    isn't purely about which parallelization approach is better; part of
    it is genuinely doing more work, and the comparison should be read
    with that in mind.",
  `drives_local-vs-network` = "Reads and writes a single sample file from
    a local temp folder (C:/) vs. directly over the network drive (Y:/), to isolate
    how much network speed (rather than processing itself) affects total
    time.",
  `workers_raw-to-processed` = "Runs our actual production processing
    step (managelidar's raw-to-processed pipeline) on all 100 sample
    files repeatedly, each time asking for a different number of
    parallel workers (50%, 75%, and 100% of the machine's logical
    cores). The most direct evidence toward the real question this
    project exists to answer: how many workers to actually use in
    production."
)

# Hand-written interpretation of each benchmark's results — not derived
# from the data, so these need to be kept in sync manually if the
# numbers change on a re-run. Shown as a callout below each benchmark's
# table(s).
benchmark_takeaways <- list(
  `drives_local-vs-network` = c(
    "Local storage is faster than network on every machine tested.",
    "PC166's network connection is much slower than the others.",
    "This only tested one file at a time — the network penalty could
     matter more once several files are read/written in parallel and
     share the same bandwidth.",
    "LB-3D2026 is the fastest machine for single-file processing."
  ),
  lasr_stages = c(
    "Stage costs vary widely across the pipeline.",
    "Ground classification is by far the heaviest stage.",
    "LB-3D2026 is the fastest machine for single-file processing."
  ),
  `lidr-lasr_ground-intensity` = c(
    "The lasR pipeline (with a custom R callback) is much faster than
     the equivalent lidR pipeline.",
    "LB-3D2026 is the fastest machine."
  ),
  parallel_strategies = c(
    "mirai is the fastest parallelization strategy when no buffer
     reading is needed (see the lasr_internal caveat above).",
    "PC026 is the fastest machine, using half of its many cores."
  ),
  tools_copc = c(
    "lasR is faster than PDAL; LAStools is faster than lasR.",
    "A new/updated lasR COPC writer is reportedly in development, which
     may change this ranking.",
    "If COPC conversion is just one step within a larger lasR pipeline,
     lasR is probably still the fastest choice overall in practice,
     since it avoids the extra read/write round-trip a separate
     conversion step would need.",
    "LB-3D2026 is the fastest machine for single-file processing."
  ),
  tools_ground = c(
    "LAStools is the fastest tool overall. (Not sure whether it can make use of more CPU)",
    "Among the R-based options, lasR's PTD algorithm is the fastest —
     and gives good classification results too.",
    "LB-3D2026 is the fastest machine."
  ),
  tools_normals = c(
    "lasR is faster than PDAL."
  ),
  tools_rw = c(
    "LAStools is the fastest tool. (Not sure whether it can make use of more CPU)"
  ),
  `workers_raw-to-processed` = c(
    "On the machines that completed all 100 files at every worker count
     (LB-3D2026, pc069, PC166), using 100% of cores is consistently
     faster than 50-75% — 12-33% faster than at 50%.",
    "PC026 only completed all 100 files at 50% of cores; at 75% and
     100% most files failed (13/100 and 16/100 completed, see the
     footnotes on the table). Its fast times there are an artifact of
     that failure, not a real speed advantage — this points to a
     stability or resource-contention issue at very high concurrency on
     that machine that needs investigating before running production
     jobs with that many workers."
  )
)

# Prints a "Key takeaways" callout for a benchmark, if one is defined.
print_takeaways <- function(b) {
  if (!b %in% names(benchmark_takeaways)) return(invisible(NULL))
  cat('\n::: {.callout-tip title="Key takeaways"}\n\n')
  cat(paste0("- ", gsub("\\s+", " ", benchmark_takeaways[[b]]), collapse = "\n"))
  cat("\n\n:::\n\n")
}

# Bolds+shades the fastest machine's cell in each row of a wide (id x
# machine) gt table. Kept separate from pivot_wide() (R/report-helpers.R)
# since it's gt-specific styling, not data reshaping.
highlight_fastest <- function(tbl, wide, machines) {
  row_min <- apply(wide[machines], 1, function(x) {
    if (all(is.na(x))) NA_real_ else min(x, na.rm = TRUE)
  })
  for (m in machines) {
    fastest_rows <- which(!is.na(wide[[m]]) & wide[[m]] == row_min)
    if (length(fastest_rows) > 0) {
      tbl <- tbl |> tab_style(
        style = list(cell_text(weight = "bold"), cell_fill(color = "#e8f4ea")),
        locations = cells_body(columns = tidyselect::all_of(m), rows = fastest_rows)
      )
    }
  }
  tbl
}

# Bolds+shades the single fastest cell across the whole wide (id x
# machine) table, on top of any per-row highlight_fastest() styling —
# a distinct color so "fastest overall" reads differently from
# "fastest machine in this row".
highlight_overall_fastest <- function(tbl, wide, machines) {
  mat <- as.matrix(wide[machines])
  if (all(is.na(mat))) return(tbl)
  min_val <- min(mat, na.rm = TRUE)
  for (m in machines) {
    rows <- which(!is.na(wide[[m]]) & wide[[m]] == min_val)
    if (length(rows) > 0) {
      tbl <- tbl |> tab_style(
        style = list(
          cell_text(weight = "bold"),
          cell_fill(color = "#ffe9a8"),
          cell_borders(sides = "all", color = "#c9a227", weight = px(2))
        ),
        locations = cells_body(columns = tidyselect::all_of(m), rows = rows)
      )
    }
  }
  tbl
}

# Named vector of file count per machine for a benchmark that ran against
# a full corpus (filesetinfo present), NULL if this benchmark only ever
# ran against a single representative file. Used to turn a total
# wall-clock time into an average time per file for corpus-wide
# benchmarks (parallel_strategies, workers_raw-to-processed) — the more
# comparable, extrapolation-ready number for planning full-corpus runs.
n_files_by_machine <- function(meta, benchmark) {
  entries <- Filter(function(m) m$benchmark == benchmark && !is.null(m$filesetinfo), meta)
  if (length(entries) == 0) return(NULL)
  setNames(
    vapply(entries, function(m) m$filesetinfo$n_files, numeric(1)),
    vapply(entries, function(m) m$nodename, character(1))
  )
}

# Divides each machine column in a wide (id x machine) table by that
# machine's file count, turning total wall-clock time into average time
# per file. Machines missing from `nfiles` become NA.
per_file_wide <- function(wide, machines, nfiles) {
  for (m in machines) {
    wide[[m]] <- if (m %in% names(nfiles) && nfiles[[m]] > 0) wide[[m]] / nfiles[[m]] else NA_real_
  }
  wide
}

# Named vector of logical core count per machine for a benchmark, from
# systeminfo — used to express a swept `workers` value as % of that
# machine's cores instead of an absolute count, since the same
# percentage means a different absolute worker count on every machine.
cores_by_machine <- function(meta, benchmark) {
  entries <- Filter(function(m) m$benchmark == benchmark && !is.null(m$systeminfo), meta)
  if (length(entries) == 0) return(NULL)
  setNames(
    vapply(entries, function(m) m$systeminfo$cores, numeric(1)),
    vapply(entries, function(m) m$nodename, character(1))
  )
}
```

Re-render after adding new results: `quarto render report/benchmark-report.qmd`.

## Machines

```{r machines-table}
machine_rows <- lapply(results$meta, function(b) {
  si <- b$systeminfo
  data.frame(
    nodename = b$nodename,
    cpu      = if (!is.null(si)) si$cpu else NA,
    cores    = if (!is.null(si)) si$cores else NA,
    ram_gb   = if (!is.null(si)) round(as.numeric(si$ram) / 1024^3, 1) else NA
  )
})
machines <- unique(do.call(rbind, machine_rows))

gt(machines) |>
  cols_label(
    nodename = "machine", cpu = "CPU", cores = "cores", ram_gb = "RAM (GB)"
  ) |>
  fmt_number(columns = ram_gb, decimals = 1) |>
  opt_row_striping() |>
  tab_options(table.font.size = "13px", data_row.padding = px(4))
```

## Test data

All benchmarks run against a fixed random sample of 100 real ALS tiles
(~4.5GB total, seed 42) from Lower Saxony's 2016 statewide ALS campaign.
Most tool/stage comparisons use one representative file from that
sample; the corpus-wide benchmarks (parallelization strategies,
worker-count scaling) use the full 100 files. See the README for how
the sample was chosen and how to regenerate it.

## Package versions

```{r versions-table}
#| results: asis
version_rows <- lapply(results$meta, function(b) {
  pv <- b$pkgversions
  data.frame(
    benchmark = display_title(b$benchmark),
    nodename  = b$nodename,
    r_version = if (!is.null(pv)) pv$r else NA,
    lasR      = if (!is.null(pv)) pv$packages$lasR else NA,
    lidR      = if (!is.null(pv)) pv$packages$lidR else NA
  )
})

tbl <- gt(do.call(rbind, version_rows)) |>
  cols_label(
    benchmark = "benchmark", nodename = "machine",
    r_version = "R", lasR = "lasR", lidR = "lidR"
  ) |>
  opt_row_striping() |>
  tab_options(table.font.size = "13px", data_row.padding = px(4))

cat("<details>\n<summary>R/lasR/lidR versions per benchmark run (this repo doesn't pin a lockfile, so versions can drift over time — see README)</summary>\n\n")
print(tbl)
cat("\n\n</details>\n\n")
```

## Named comparisons (tool/stage benchmarks)

```{r named-results}
#| results: asis
if (!is.null(results$named)) {
  for (b in unique(results$named$benchmark)) {
    cat("###", display_title(b), "\n\n")
    if (b %in% names(benchmark_context)) cat(benchmark_context[[b]], "\n\n")

    d <- results$named[results$named$benchmark == b, c("nodename", "expression", "elapsed")]
    d <- mask_missing_external_tools(d, results$meta, b)
    d$elapsed <- na_if_unavailable(d$elapsed)
    machines <- unique(d$nodename)
    # One row per method, one column per machine — reading down a column
    # compares methods on one machine, reading across a row compares
    # machines on one method, without the two mixing into one ranking.
    wide <- pivot_wide(d, "expression", "nodename", "elapsed")

    # Benchmarks that ran against the full corpus (not just one
    # representative file) report average time per file instead of total
    # wall-clock time — the more comparable, extrapolation-ready number.
    # Single-file benchmarks just report the (already per-file) time.
    nfiles <- n_files_by_machine(results$meta, b)
    display <- if (!is.null(nfiles)) per_file_wide(wide, machines, nfiles) else wide

    if (!is.null(nfiles)) cat("Average time per file (s):\n\n") else cat("Time (s):\n\n")

    # lasr_stages gets a "Total" row summing every stage's time — useful
    # as a rough total-pipeline-cost signal, though note some stages
    # overlap (dtm/dsm/normalize internally reuse triangulate/rasterize,
    # and the four classify_with_* rows are alternative methods, not
    # steps of the same pipeline), so it's not a literal single run time.
    is_lasr_stages <- b == "lasr_stages"
    tbl <- if (is_lasr_stages) {
      gt(display, rowname_col = "expression") |> tab_stubhead(label = "method")
    } else {
      gt(display) |> cols_label(expression = "method")
    }
    tbl <- tbl |>
      fmt_number(columns = tidyselect::all_of(machines), decimals = 1) |>
      sub_missing(columns = tidyselect::all_of(machines), missing_text = "not available") |>
      opt_row_striping() |>
      tab_options(table.font.size = "13px", data_row.padding = px(4))

    if (is_lasr_stages) {
      tbl <- tbl |> grand_summary_rows(
        columns = tidyselect::all_of(machines),
        fns = list(label = "Total", id = "total", fn = "sum"),
        fmt = ~ fmt_number(., decimals = 1),
        missing_text = "not available"
      )
    }

    # Skip fastest-machine highlighting on the average-time-per-file
    # table (parallel_strategies) to keep it simpler — only single-file
    # tables get the per-row fastest-machine marker.
    if (is.null(nfiles) && length(machines) > 1) {
      tbl <- tbl |>
        tab_source_note("bold = fastest machine for that method") |>
        highlight_fastest(display, machines)
    }

    print(tbl)
    cat("\n\n")
    print_takeaways(b)
  }
} else {
  cat("No named-comparison results found under data/benchmarks/.\n")
}
```

## Parameter sweeps (e.g. worker-count scaling)

```{r sweep-results}
#| results: asis
if (!is.null(results$sweep)) {
  for (b in unique(results$sweep$benchmark)) {
    cat("###", display_title(b), "\n\n")
    if (b %in% names(benchmark_context)) cat(benchmark_context[[b]], "\n\n")

    d <- results$sweep[results$sweep$benchmark == b, ]
    param <- setdiff(names(d), c("seconds", "nodename", "benchmark", "n_ok", "n_total"))[1]
    has_success_info <- all(c("n_ok", "n_total") %in% names(d))
    d <- d[, c("nodename", param, "seconds", if (has_success_info) c("n_ok", "n_total"))]
    d$seconds <- na_if_unavailable(d$seconds)

    # Show `workers` as % of each machine's own core count rather than an
    # absolute number — machines have different core counts, so the same
    # absolute worker count means something different on each, while the
    # same percentage lines up as the same row across machines.
    is_pct <- FALSE
    if (param == "workers") {
      cores <- cores_by_machine(results$meta, b)
      if (!is.null(cores) && all(d$nodename %in% names(cores))) {
        d[[param]] <- round(d[[param]] / cores[d$nodename] * 100)
        names(d)[names(d) == param] <- "% of cores"
        param <- "% of cores"
        is_pct <- TRUE
      }
    }

    machines <- unique(d$nodename)
    # One row per parameter value, one column per machine — see the
    # named-comparisons loop above for why this beats one long table.
    wide <- pivot_wide(d, param, "nodename", "seconds")
    wide <- wide[order(wide[[param]]), ]

    # Report average time per file when the run covered a known corpus
    # size (always true today, kept conditional for robustness) — see
    # the named-comparisons loop above for why.
    nfiles <- n_files_by_machine(results$meta, b)
    display <- if (!is.null(nfiles)) per_file_wide(wide, machines, nfiles) else wide

    if (!is.null(nfiles)) cat("Average time per file (s):\n\n") else cat("Time (s):\n\n")

    tbl <- gt(display) |>
      fmt_number(columns = tidyselect::all_of(machines), decimals = 1) |>
      sub_missing(columns = tidyselect::all_of(machines), missing_text = "not available") |>
      opt_row_striping() |>
      tab_options(table.font.size = "13px", data_row.padding = px(4))

    if (is_pct) {
      tbl <- tbl |> fmt_number(columns = tidyselect::all_of(param), decimals = 0, pattern = "{x}%")
    }

    # Show each machine's core count in its column header — useful here
    # since "% of cores" only means something once you know how many
    # cores that percentage is relative to on each machine.
    col_cores <- cores_by_machine(results$meta, b)
    if (!is.null(col_cores)) {
      machine_labels <- setNames(
        lapply(machines, function(m) {
          if (m %in% names(col_cores)) paste0(m, " (", col_cores[[m]], ")") else m
        }),
        machines
      )
      tbl <- tbl |> cols_label(.list = machine_labels)
    }

    # A run that only completed a fraction of its files can look
    # deceptively fast — exclude those cells from "fastest" ranking (a
    # separate copy, so the real number still prints) and footnote them
    # on the table itself so the reader isn't misled by the raw time.
    ranking_display <- display
    if (has_success_info) {
      incomplete <- d[!is.na(d$n_ok) & !is.na(d$n_total) & d$n_ok < d$n_total, ]
      for (i in seq_len(nrow(incomplete))) {
        row_idx <- which(display[[param]] == incomplete[[param]][i])
        m <- incomplete$nodename[i]
        if (length(row_idx) == 1 && m %in% names(ranking_display)) {
          ranking_display[row_idx, m] <- NA_real_
          tbl <- tbl |> tab_footnote(
            footnote = sprintf("only %d/%d files completed successfully — time not comparable",
                                incomplete$n_ok[i], incomplete$n_total[i]),
            locations = cells_body(columns = tidyselect::all_of(m), rows = row_idx)
          )
        }
      }
    }

    # Skip per-row (green) highlighting on the average-time-per-file
    # table to keep it simpler — the single fastest-overall (gold)
    # marker is enough there.
    notes <- character(0)
    if (is.null(nfiles) && length(machines) > 1) {
      tbl <- tbl |> highlight_fastest(ranking_display, machines)
      notes <- c(notes, "green = fastest machine for that parameter value")
    }
    tbl <- tbl |> highlight_overall_fastest(ranking_display, machines)
    notes <- c(notes, "gold = fastest overall")
    tbl <- tbl |> tab_source_note(paste(notes, collapse = "; "))

    print(tbl)
    cat("\n\n")
    print_takeaways(b)
  }
} else {
  cat("No parameter-sweep results found under data/benchmarks/ yet (e.g. run benchmarks/workers_raw-to-processed.R).\n")
}
```
