---
title: "ALS Processing Benchmark Report"
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: false
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load}
source("R/report-helpers.R")

results <- load_all_benchmarks()
```

This report is generated entirely from `data/benchmarks/**/*.RDS` — it
discovers every machine folder and benchmark result present, so it never
goes stale the way a hardcoded script would. Re-render after adding new
results with `quarto render report/benchmark-report.qmd`.

## System & package versions

Every timing below is traceable to the machine and tool versions that
produced it.

```{r sysinfo-table}
sysinfo_rows <- lapply(results$meta, function(b) {
  si <- b$systeminfo
  pv <- b$pkgversions
  data.frame(
    benchmark = b$benchmark,
    nodename  = b$nodename,
    cpu       = if (!is.null(si)) si$cpu else NA,
    cores     = if (!is.null(si)) si$cores else NA,
    ram_gb    = if (!is.null(si)) round(as.numeric(si$ram) / 1024^3, 1) else NA,
    r_version = if (!is.null(pv)) pv$r else NA,
    lasR      = if (!is.null(pv)) pv$packages$lasR else NA,
    lidR      = if (!is.null(pv)) pv$packages$lidR else NA
  )
})

knitr::kable(do.call(rbind, sysinfo_rows), row.names = FALSE)
```

## Named comparisons (tool/stage benchmarks)

One table and chart per benchmark type, across every machine that has
results for it.

```{r named-results}
#| results: asis
if (!is.null(results$named)) {
  for (b in unique(results$named$benchmark)) {
    cat("###", b, "\n\n")
    d <- results$named[results$named$benchmark == b, ]
    print(knitr::kable(d[, c("nodename", "expression", "elapsed")], row.names = FALSE))
    cat("\n\n")

    p <- ggplot2::ggplot(d, ggplot2::aes(x = expression, y = elapsed, fill = nodename)) +
      ggplot2::geom_col(position = "dodge") +
      ggplot2::coord_flip() +
      ggplot2::labs(x = NULL, y = "elapsed (s)", title = b) +
      ggplot2::theme_minimal()
    print(p)
    cat("\n\n")
  }
} else {
  cat("No named-comparison results found under data/benchmarks/.\n")
}
```

## Parameter sweeps (e.g. worker-count scaling)

```{r sweep-results}
#| results: asis
if (!is.null(results$sweep)) {
  for (b in unique(results$sweep$benchmark)) {
    cat("###", b, "\n\n")
    d <- results$sweep[results$sweep$benchmark == b, ]
    param <- setdiff(names(d), c("seconds", "nodename", "benchmark"))[1]

    p <- ggplot2::ggplot(d, ggplot2::aes(x = .data[[param]], y = seconds, color = nodename)) +
      ggplot2::geom_line() +
      ggplot2::geom_point() +
      ggplot2::labs(x = param, y = "seconds", title = b) +
      ggplot2::theme_minimal()
    print(p)
    cat("\n\n")
  }
} else {
  cat("No parameter-sweep results found under data/benchmarks/ yet (e.g. run benchmarks/workers_raw-to-processed.R).\n")
}
```
