---
title: "ALS Processing Benchmark Report"
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: false
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load}
#| include: false
source("R/report-helpers.R")
library(gt)

results <- load_all_benchmarks()
```

This report is generated entirely from `data/benchmarks/**/*.RDS` — it
discovers every machine folder and benchmark result present, so it never
goes stale the way a hardcoded script would. Re-render after adding new
results with `quarto render report/benchmark-report.qmd`.

## System & package versions

Every timing below is traceable to the machine and tool versions that
produced it.

```{r sysinfo-table}
sysinfo_rows <- lapply(results$meta, function(b) {
  si <- b$systeminfo
  pv <- b$pkgversions
  data.frame(
    benchmark = b$benchmark,
    nodename  = b$nodename,
    cpu       = if (!is.null(si)) si$cpu else NA,
    cores     = if (!is.null(si)) si$cores else NA,
    ram_gb    = if (!is.null(si)) round(as.numeric(si$ram) / 1024^3, 1) else NA,
    r_version = if (!is.null(pv)) pv$r else NA,
    lasR      = if (!is.null(pv)) pv$packages$lasR else NA,
    lidR      = if (!is.null(pv)) pv$packages$lidR else NA
  )
})

gt(do.call(rbind, sysinfo_rows)) |>
  cols_label(
    benchmark = "benchmark", nodename = "machine", cpu = "CPU",
    cores = "cores", ram_gb = "RAM (GB)", r_version = "R",
    lasR = "lasR", lidR = "lidR"
  ) |>
  fmt_number(columns = ram_gb, decimals = 1) |>
  opt_row_striping() |>
  tab_options(table.font.size = "13px", data_row.padding = px(4))
```

## Named comparisons (tool/stage benchmarks)

One minimal table per benchmark type, across every machine that has
results for it. Relative time is scaled against the fastest result in
each table (1.00 = fastest).

```{r named-results}
#| results: asis
if (!is.null(results$named)) {
  for (b in unique(results$named$benchmark)) {
    cat("###", b, "\n\n")
    d <- results$named[results$named$benchmark == b, c("nodename", "expression", "elapsed")]
    d <- add_relative_time(d, "elapsed")

    tbl <- gt(d) |>
      cols_label(
        nodename = "machine", expression = "method",
        elapsed = "time (s)", relative = "relative"
      ) |>
      fmt_number(columns = elapsed, decimals = 2) |>
      fmt_number(columns = relative, decimals = 2, pattern = "{x}x") |>
      sub_missing(columns = c(elapsed, relative), missing_text = "not available") |>
      data_color(
        columns = relative,
        method = "numeric",
        palette = c("#e8f4ea", "#f9e6c8"),
        domain = c(1, max(d$relative, na.rm = TRUE))
      ) |>
      opt_row_striping() |>
      tab_options(table.font.size = "13px", data_row.padding = px(4))

    print(tbl)
    cat("\n\n")
  }
} else {
  cat("No named-comparison results found under data/benchmarks/.\n")
}
```

## Parameter sweeps (e.g. worker-count scaling)

Relative time is scaled against the fastest parameter value in each
table (1.00 = fastest).

```{r sweep-results}
#| results: asis
if (!is.null(results$sweep)) {
  for (b in unique(results$sweep$benchmark)) {
    cat("###", b, "\n\n")
    d <- results$sweep[results$sweep$benchmark == b, ]
    param <- setdiff(names(d), c("seconds", "nodename", "benchmark"))[1]
    d <- d[, c("nodename", param, "seconds")]
    d <- d[order(d[[param]]), ]
    d <- add_relative_time(d, "seconds", reorder = FALSE)

    tbl <- gt(d) |>
      cols_label(
        nodename = "machine", seconds = "time (s)", relative = "relative"
      ) |>
      fmt_number(columns = seconds, decimals = 2) |>
      fmt_number(columns = relative, decimals = 2, pattern = "{x}x") |>
      sub_missing(columns = c(seconds, relative), missing_text = "not available") |>
      data_color(
        columns = relative,
        method = "numeric",
        palette = c("#e8f4ea", "#f9e6c8"),
        domain = c(1, max(d$relative, na.rm = TRUE))
      ) |>
      opt_row_striping() |>
      tab_options(table.font.size = "13px", data_row.padding = px(4))

    print(tbl)
    cat("\n\n")
  }
} else {
  cat("No parameter-sweep results found under data/benchmarks/ yet (e.g. run benchmarks/workers_raw-to-processed.R).\n")
}
```
